---
title: "Homework 2"
author: "Amy Atwater"
date: "October 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Homework 2, Module 8 ###

```{r}
f <- file.choose()
d <- read.csv(f, sep=",", header = TRUE)
head(d)
summary(d)

### 1. ###

height = d$height
mean(height)
x<-d$height
stdv<- function(x){sqrt(sum((x - mean(x))^2)/length(x))}
### Had to make a function to calculate population standard deviations, as the built in sd() function is for sample populations, and divides the numerator by the incorrect value. x-1 instead of just x. ###
stdv(x)

weight = d$weight
mean(weight)
stdv(weight)

age = d$age
mean(age)
stdv(age)

zom = d$zombies_killed
mean(zom)
stdv(zom)

edu = d$years_of_education
mean(edu)
stdv(edu)

### 2. ###

boxplot(height ~ gender, d)
library(ggplot2)
p <- ggplot(data = d, aes(x = gender, y = height))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p

boxplot(weight ~ gender, d)
p <- ggplot(data = d, aes(x = gender, y = weight))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p

boxplot(age ~ gender, d)
p <- ggplot(data = d, aes(x = gender, y = age))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p

boxplot(zom ~ gender, d)
p <- ggplot(data = d, aes(x = gender, y = zom))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("Number of Zombies Killed")
p

boxplot(edu~ gender, d)
p <- ggplot(data = d, aes(x = gender, y = edu))
p <- p + geom_boxplot()
p <- p + theme(axis.text.x = element_text(angle = 90))
p <- p + ylab("Years of Education")
p

### 3. ###

p <- ggplot(data = d, aes(x = age, y = height, color = factor(gender)))
p <- p + xlab("Age") + ylab("Height")
p <- p + geom_point()
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p

p <- ggplot(data = d, aes(x = age, y = weight, color = factor(gender)))
p <- p + xlab("Age") + ylab("Weight")
p <- p + geom_point()
p <- p + theme(legend.position = "bottom", legend.title = element_blank())
p

### there seems to be more of a relationship between height and age than there is between weight and age. There is a positive trend in both plots, greater heights and weights seen in the older ages. This is only really realistic when you realize the upper limit of age is 30. ### 

### 4. ###

hist(x, freq = FALSE, col = "purple", main = "Plot 1", xlab = "Height", ylab = "Frequency", ylim = c(0, 0.15))

qqnorm(x, main = "Height QQ plot")
qqline(x, col = "gold")

hist(weight, freq = FALSE, col = "red", main = "Plot 1", xlab = "Weight", ylab = "Frequency", ylim = c(0, 0.04))

qqnorm(weight, main = "Weight QQ plot")
qqline(weight, col = "purple")

hist(age, freq = FALSE, col = "green", main = "Plot 1", xlab = "Age", ylab = "Frequency", ylim = c(0, 0.15))

qqnorm(age, main = "Age QQ plot")
qqline(age, col = "hot pink")

hist(zom, freq = FALSE, col = "gold", main = "Plot 1", xlab = "Zombies Killed", ylab = "Frequency", ylim = c(0, 0.27))

qqnorm(zom, main = "Zombies Killed QQ plot")
qqline(zom, col = "blue")

hist(edu, freq = FALSE, col = "blue", main = "Plot 1", xlab = "Years of Education", ylab = "Frequency", ylim = c(0, 0.25))

qqnorm(x, main = "Years of Education QQ plot")
qqline(x, col = "brown")

### The normally distributed variables include height, weight and age. Number of zombies killed and years of education do not show a normal distribution. They appear to be more of a Poisson distribution, as they are skewed to one side, with a much longer tail on the other side. ###

### 5 ###
s <- sample(d, size = 30, replace = TRUE)
subset <- d[sample(1: nrow(d), 30, replace, FALSE

mysample <- d[sample(1:nrow(d),30,replace=FALSE),]
mysample

height = mysample$height
sd(height)
mh <- mean(height)
var(height)
### standard error ###
sem <- sd(height)/sqrt(length(height))
sem
sd(height)/sqrt(length(height))
SE1 <- function(x) {sqrt(var(x)/length(x))}
SE1(height)

lower <- mh - qnorm(1 - 0.05/2) * sem
upper <- mh + qnorm(1 - 0.05/2) * sem
ci <- c(lower, upper)
ci


weight = mysample$weight
sd(weight)
mean(weight)
var(weight)
### standard error ###
sd(weight)/sqrt(30)
SE1(weight)
sem2 <- sd(weight)/sqrt(length(weight))
lower <- mh - qnorm(1 - 0.05/2) * sem2
upper <- mh + qnorm(1 - 0.05/2) * sem2
ci <- c(lower, upper)
ci

age = mysample$age
sd(age)
mean(age)
### standard error ###
SE1(age)
sem3 <- sd(age)/sqrt(length(age))
lower <- mh - qnorm(1 - 0.05/2) * sem3
upper <- mh + qnorm(1 - 0.05/2) * sem3
ci <- c(lower, upper)
ci

zom = mysample$zombies_killed
mean(zom)
sd(zom)
### standard error ###
SE1(zom)

### calculating confidence intervals by bootstrapping as it does not presume the underlying distribution from which the random variable is drawn ###
set <- NULL
n <- 15
for (i in 1:10000) {set[i] <- mean(sample(zom, n, replace = TRUE))}
quantile(set)
quantile(set, c(0.05, 0.95))

edu = mysample$years_of_education
mean(edu)
sd(edu)
### standard error ###
SE1(edu)
### calculating confidence intervals by bootstrapping ###
set <- NULL
n <- 15
for (i in 1:10000) {set[i] <- mean(sample(edu, n, replace = TRUE))}
quantile(set)
quantile(set, c(0.05, 0.95))


### 6. ###

k <- 99
n <- 30
s <- NULL
for (i in 1:k) s[[i]] <-d[sample(1:nrow(d),30,replace=FALSE),]

### height

m <- NULL

for (i in 1:k) {m[i] <- mean(s[[i]]$height)}
mean(m)
h_mean<-c(m,mean(mysample$height))
h_mean

stdev <- NULL
for(i in 1:k){stdev[i]<-sd(s[[i]]$height)}

h_sd<-c(stdev,sd(mysample$height))
qqnorm(h_sd)
hist(h_mean)
mean(h_mean)
sd(h_sd)

### weight ###
w <- NULL

for (i in 1:k) {w[i] <- mean(s[[i]]$height)}
mean(w)
w_mean<-c(w,mean(mysample$weight))
w_mean

stdev <- NULL
for(i in 1:k){stdev[i]<-sd(s[[i]]$weight)}

w_sd<-c(stdev,sd(mysample$weight))
qqnorm(w_sd)
hist(w_mean)
mean(w_mean)
sd(w_sd)

### Age ###
a <- NULL

for (i in 1:k) {a[i] <- mean(s[[i]]$age)}
mean(a)
a_mean<-c(a,mean(mysample$age))
a_mean

stdev <- NULL
for(i in 1:k){stdev[i]<-sd(s[[i]]$height)}

a_sd<-c(stdev,sd(mysample$age))
qqnorm(a_sd)
hist(a_mean)
mean(a_mean)
sd(a_sd)


### Zombies Killed ###
z <- NULL

for (i in 1:k) {z[i] <- mean(s[[i]]$zom)}
mean(z)
z_mean<-c(z,mean(mysample$zom))
z_mean

stdev <- NULL
for(i in 1:k){stdev[i]<-sd(s[[i]]$height)}

z_sd<-c(stdev,sd(mysample$zom))
qqnorm(z_sd)
hist(z_mean)
mean(z_mean)
sd(z_sd)

### Years of Education ###
e <- NULL

for (i in 1:k) {e[i] <- mean(s[[i]]$edu)}
mean(e)
e_mean<-c(e,mean(mysample$edu))
e_mean

stdev <- NULL
for(i in 1:k){stdev[i]<-sd(s[[i]]$height)}

e_sd<-c(stdev,sd(mysample$edu))
qqnorm(e_sd)
hist(e_mean)
mean(e_mean)
sd(e_sd)

#### The standard deviations calculated in part 9 are less than the standard error values calculated in part 5, greater sampling reduces the amount of deviation. The distributions of all the variables now look normally distributed, even zombies killed and years of education, which previously seemed skewed. Repeated sampling allows for a more normal distribution of data. ###

